Supporting Information

### 2015-08-28 ms.Rmd
fixed gene name error in zhang in situ db. need to re-run table generator

---
title: Genome-wide changes in expression during inflorescence architecture development
  in rice
csl: csl/the-plant-journal.csl
header-includes: \usepackage{graphicx}
output:
  BiocStyle::html_document:
    fig_caption: yes
    force_captions: yes
    highlight: pygments
    keep_md: yes
    number_sections: yes
    dev: "CairoPDF"
  word_document:
    fig_caption: yes
    reference_docx: csl/wordStyles.docx
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    number_sections: yes
bibliography: bib/test.bib
---

@@

To recover common expression patterns in genes detected in the meristem samples, read counts were transformed using the variance-stabilizing transformation (VST) included in the `DESeq2` package [@Love:2014do]. Expressed genes were ranked by variance, and standardised, VST-transformed read counts for the 25% most-variable genes were clustered using the fuzzy *c*-means algorithm implemented in the `Mfuzz` package [@Futschik:2005vo]. To determine the number of cluster cores (*c*), the minimum distance between cluster centroids, the formation of empty clusters, and PCA plots of cluster members were monitored for clusters produced with *c* values between 2 and `r maxClust` (`r I(pasteLabel("Figures S", s_figCount, "sf_mfuzzCentroids"))` `r I(pasteLabel(" and S", s_figCount, "sf_mfuzzPca"))`), leading to the use of `r as.character(english::english(c))` cluster cores (`r I(pasteLabel("Figure ", figCount, "f_mfuzzClusters"))`). There are several expression patterns descibed by the clusters: genes that increase or decrease steadily during development (clusters `r getClusterName('inc', 'inc')` and `r getClusterName('dec', 'dec')`), genes that change expression between RM and PBM and remain constant from PBM to SM (clusters `r getClusterName('inc', 'small')` and `r getClusterName('dec', 'small')`), genes that are constant between RM and PBM and then change between PBM and SM (clusters `r getClusterName('small', 'inc')` and `r getClusterName('small', 'dec')`). There were also two weaker clusters containing fewer genes, which had complex expression patterns involving changes in expression particular to the PBM (clusters `r getClusterName('inc', 'dec')` and `r getClusterName('dec', 'inc')`).

@@

Analysis of the sequenced libraries was carried out with custom `bash` and `R` (**ref**) scripts, which were arranged into a pipeline using the `ruffus` package for `python3` (**ref**). All of the code used for the analysis and to generate text, figures and tables for this report, along with the versions and paramaters of the software used, are available with revision history in a public GitHub repository at https://github.com/evoreprice/lmdPaper.

Briefly, adaptor trimming was performed with `cutadapt` (**ref**) and reads were mapped with the MSUv7 genome and annotation (**ref**) downloaded from Phytozome 10.3 (**ref**) using `rna-star` (**ref**) in 2-pass mode. The `quantMode` argument of `rna-star` was used for read counting. rRNA and tRNA contamination was estimated using `htseq-count` (**ref**) to count the number of reads that mapped to regions annotated as tRNA or rRNA on the Rap-DB (**ref**) or to regions where reads simulated with `wgsim` (**ref**) from rRNA sequences in the Plant Repeats Database (**ref**) also mapped. Differential expression analysis to calculate Log~2~ fold change (L~2~FC) values and transformed read counts was performed with the `DESeq2` package (**ref**). To determine a strict cutoff for unexpressed genes, expression values in transcripts per million (TPM) were calculated for each gene (**2 refs**). These values were compared to pseudo-expression values calculated from intergenic regions of the genome, using intervals of a similar size distrubution to those used for calculating the 'genic' expression values. For each library, the cutoff was placed at the 95^th^ percentile of the distribution of intergenic expression values. Transformed counts and L~2~FC values from expressed genes were used for soft clustering with the `Mfuzz` package (**ref**) and geneset enrichment analysis with the `gage` package (**ref**) respectively. All plots were produced with the `ggplot2` package (**ref**).

>>>>

---
title: Genome-wide changes in expression during inflorescence architecture development
  in rice
csl: csl/the-plant-journal.csl
header-includes: \usepackage{graphicx}
output:
  BiocStyle::html_document:
    fig_caption: yes
    force_captions: yes
    highlight: pygments
    keep_md: yes
    number_sections: yes
    dev: "CairoPDF"
  word_document:
    fig_caption: yes
    reference_docx: csl/wordStyles.docx
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    number_sections: yes
bibliography: bib/papersLib.bib
---

@@

To recover common expression patterns in genes detected in the meristem samples, read counts were transformed using the variance-stabilizing transformation (VST) included in the `DESeq2` package [@Love:2014do]. Expressed genes were ranked by variance, and standardised, VST-transformed read counts for the 25% most-variable genes were clustered using the fuzzy *c*-means algorithm implemented in the `Mfuzz` package [@Kumar:2007uw]. To determine the number of cluster cores (*c*), the minimum distance between cluster centroids, the formation of empty clusters, and PCA plots of cluster members were monitored for clusters produced with *c* values between 2 and `r maxClust` (`r I(pasteLabel("Figures S", s_figCount, "sf_mfuzzCentroids"))` `r I(pasteLabel(" and S", s_figCount, "sf_mfuzzPca"))`), leading to the use of `r as.character(english::english(c))` cluster cores (`r I(pasteLabel("Figure ", figCount, "f_mfuzzClusters"))`). There are several expression patterns descibed by the clusters: genes that increase or decrease steadily during development (clusters `r getClusterName('inc', 'inc')` and `r getClusterName('dec', 'dec')`), genes that change expression between RM and PBM and remain constant from PBM to SM (clusters `r getClusterName('inc', 'small')` and `r getClusterName('dec', 'small')`), genes that are constant between RM and PBM and then change between PBM and SM (clusters `r getClusterName('small', 'inc')` and `r getClusterName('small', 'dec')`). There were also two weaker clusters containing fewer genes, which had complex expression patterns involving changes in expression particular to the PBM (clusters `r getClusterName('inc', 'dec')` and `r getClusterName('dec', 'inc')`).

@@

Analysis of the sequenced libraries was carried out with custom `bash` and `R` scripts [@RALanguageandEn:2015wf], which were arranged into a pipeline using the `ruffus` package for `python3` [@Goodstadt:2010jp]. All of the code used for the analysis and to generate text, figures and tables for this report, along with the versions and paramaters of the software used, are available with revision history in a public GitHub repository at https://github.com/evoreprice/lmdPaper.

Briefly, adaptor trimming was performed with `cutadapt` [@Martin:2011eu] and reads were mapped with the MSUv7 genome and annotation [@Kawahara:2013gh; @Ouyang:2007il] downloaded from Phytozome 10.3 [@Goodstein:2012ho] using `STAR` [@Dobin:2012fg] in 2-pass mode . The `'quantMode'` argument of `STAR` was used for read counting. rRNA and tRNA contamination was estimated using `htseq-count` [@Anders:2014gf] to count the number of reads that mapped to regions annotated as tRNA or rRNA on the Rap-DB (**ref**) or to regions where reads simulated with [`wgsim`](https://github.com/lh3/wgsim) from rRNA sequences in the TIGR Plant Repeats Database [@Ouyang:2004ia] also mapped. Differential expression analysis to calculate Log~2~ fold change (L~2~FC) values and transformed read counts was performed with the `DESeq2` package [@Love:2014do]. To determine a strict cutoff for unexpressed genes, expression values in transcripts per million were calculated for each gene [@Li:2010cl; @Wagner:2012jf]. These values were compared to pseudo-expression values calculated from intergenic regions of the genome, using intervals of a similar size distrubution to those used for calculating the 'genic' expression values. For each library, the cutoff was placed at the 95^th^ percentile of the distribution of intergenic expression values. Transformed counts and L~2~FC values from expressed genes were used for soft clustering with the `Mfuzz` package [@Kumar:2007uw] and geneset enrichment analysis with the `gage` package [@Luo:2009kp] respectively. All plots were produced with the `ggplot2` package [@Wickham:2009tn].

### 2015-08-28 si.Rmd
fixed gene name error in zhang in situ db. need to re-run table generator

---
title: "Supporting Information"
csl: csl/the-plant-journal.csl
header-includes: \usepackage{graphicx}
output:
  BiocStyle::html_document:
    fig_caption: yes
    force_captions: yes
    highlight: pygments
    keep_md: yes
    number_sections: yes
    dev: "CairoPDF"
  word_document:
    fig_caption: yes
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    number_sections: yes
bibliography: bib/test.bib
---

@@

>>>>

---
title: "Supporting Information"
csl: csl/the-plant-journal.csl
header-includes: \usepackage{graphicx}
output:
  BiocStyle::html_document:
    fig_caption: yes
    force_captions: yes
    highlight: pygments
    keep_md: yes
    number_sections: yes
    dev: "CairoPDF"
  word_document:
    fig_caption: yes
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    number_sections: yes
bibliography: bib/papersLib.bib
---

@@

# References for SI (citations will be included in *in situ* table)

`r pander::pander(list("85: @Li:2011es", "118: @Ren:2013jv", "129: @Suzaki:2004ib", 
    "21: @Chu:2006fw", "82: @Lee:2012hj", "154: @Yoshida:2013ff", "148: @Xue:2008ki", 
    "3: @Ashikari:2005eg", "150: @Yan:2011hw", "86: @Li:2013iq", "153: @Yoshida:2012fg", 
    "78: @Kurakawa:2007go", "82: @Lee:2012hj", "83: @Lee:2007cj", "48: @Gao:2010iz", 
    "59: @IkedaKawakatsu:2012co", "54: @Horigome:2009gt", "83: @Lee:2007cj", 
    "68: @Jiao:2010ft", "100: @Miura:2010it"))`
